---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
devtools::load_all()
deletaion_log <- cleaningtools::cleaningtools_cleaning_log |> dplyr::filter(change_type == "remove_survey")
cleaning_log2 <- cleaningtools::cleaningtools_cleaning_log |> dplyr::filter(change_type != "remove_survey")
options(scipen = 999)
```

## {cleaningtools}

<!-- badges: start -->
[![Contributor Covenant](https://img.shields.io/badge/Contributor%20Covenant-2.1-4baaaa.svg)](code_of_conduct.md)
[![R-CMD-check](https://github.com/impact-initiatives/cleaningtools/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/impact-initiatives/cleaningtools/actions/workflows/R-CMD-check.yaml)
[![codecov](https://codecov.io/gh/impact-initiatives/cleaningtools/branch/master/graph/badge.svg?token=SOH3NGXQDU)](https://codecov.io/gh/impact-initiatives/cleaningtools)

<!-- badges: end -->



The `cleaningtools` package focuses on cleaning, and has three components: 

**1. Check**, which includes a set of functions that flag values, such as check_outliers and check_logical.  

**2. Create**, which includes a set of functions to create different items for use in cleaning, such as the cleaning log from the checks, clean data, and enumerator performance.  

**3. Review**, which includes a set of functions to review the cleaning.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` {r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("impact-initiatives/cleaningtools")
```



## Code of Conduct

Please note that the cleaningtools project is released with a [Contributor Code of Conduct](https://impact-initiatives.github.io/cleaningtools/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.


