---
output: github_document
---



```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


## cleaningtools

<!-- badges: start -->
[![R-CMD-check](https://github.com/impact-initiatives/cleaningtools/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/impact-initiatives/cleaningtools/actions/workflows/R-CMD-check.yaml)
[![codecov](https://codecov.io/gh/impact-initiatives/cleaningtools/branch/master/graph/badge.svg?token=SOH3NGXQDU)](https://codecov.io/gh/impact-initiatives/cleaningtools)

<!-- badges: end -->


## Overview

The `cleaningtools` package focuses on cleaning, and has three components: <p>
**1. Check**, which includes a set of functions that flag values, such as check_outliers and check_logical.
<br> **2. Create**, which includes a set of functions to create different items for use in cleaning, such as the cleaning log from the checks, clean data, and enumerator performance.
<br> **3. Review**, which includes a set of functions to review the cleaning, such as reviewing the cleaning.


## Installation
You can install the development version from [GitHub](https://github.com/) with:

``` {r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("impact-initiatives/cleaningtools")
```


#### Example:: Creating cleaning log from raw data and clean data
`create_cleaning_log` function takes raw data and clean data as inputs and its identify any changes bwtween them and finally provide the output as claning log format.
``` {r, eval=FALSE}

cleaning_log <- create_cleaning_log(raw_data = raw_data,raw_data_uuid = "X_uuid",
                                    clean_data = clean_data,clean_data_uuid = "X_uuid",
                                    check_for_deletion_log = T,check_for_variable_name = T ))

```

#### Example:: Comparing cleaning log with clean data and raw data
`compare_cl_with_datasets` function takes raw data, clean data and cleaning log as inputs, and it first creates the cleaning log by comparing raw data and clean data, then compares it with the user-provided cleaning log. Finally, flagged the discrepancies between them (if any). 
``` {r, eval=FALSE}

comapre_cl_with_datasets(raw_data = raw_data,raw_data_uuid = "X_uuid",
                         clean_data = clean_data,clean_data_uuid = "X_uuid",
                         cleaning_log = cleaning_log2,cleaning_log_uuid = "X_uuid",
                         cleaning_log_question_name = "questions",
                         cleaning_log_new_value =  "new_value",cleaning_log_old_value = "old_value",
                         deletion_log = deletaion_log,deletion_log_uuid = "X_uuid",
                         check_for_deletion_log =T,check_for_variable_name = T))
```

