---
output: github_document
---



```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
devtools::load_all()
deletaion_log <- cleaning_log |> filter(change_type == "remove_survey")
cleaning_log2 <- cleaning_log |> filter(change_type != "remove_survey")
options(scipen = 999)

```


## cleaningtools

<!-- badges: start -->
[![R-CMD-check](https://github.com/impact-initiatives/cleaningtools/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/impact-initiatives/cleaningtools/actions/workflows/R-CMD-check.yaml)
[![codecov](https://codecov.io/gh/impact-initiatives/cleaningtools/branch/master/graph/badge.svg?token=SOH3NGXQDU)](https://codecov.io/gh/impact-initiatives/cleaningtools)

<!-- badges: end -->


## Overview

The `cleaningtools` package focuses on cleaning, and has three components: <p>
**1. Check**, which includes a set of functions that flag values, such as check_outliers and check_logical.
<br> **2. Create**, which includes a set of functions to create different items for use in cleaning, such as the cleaning log from the checks, clean data, and enumerator performance.
<br> **3. Review**, which includes a set of functions to review the cleaning, such as reviewing the cleaning.


## Installation
You can install the development version from [GitHub](https://github.com/) with:

``` {r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("impact-initiatives/cleaningtools")

```

## Examples

### Example:: Check for duplicates
```{r}
library(cleaningtools)
testdata <- data.frame(uuid = c(letters[1:4], "a", "b", "c"),
                       col_a = runif(7),
                       col_b = runif(7)) %>%
  dplyr::rename(`_uuid` = uuid)
testdata
check_duplicate(testdata)
```

Or you can check duplicate for a specific variable or combination of variables.
```{r}
testdata2 <- data.frame(
  uuid = letters[c(1:7)],
  village = paste("village", c(1:3,1:3,4)),
  ki_identifier = paste0("xx_", c(1:5,3,4))
  ) %>%
  dplyr::rename(`_uuid` = uuid)
check_duplicate(testdata2, .col_to_check = "village")
check_duplicate(testdata2, .col_to_check = c("village", "ki_identifier"), uuid = "_uuid")
```

### Example:: Creating cleaning log from raw data and clean data
`create_cleaning_log` function takes raw data and clean data as inputs and its identify any changes between them and finally provide the output as cleaning log format.
``` {r, eval=T, message=FALSE}
cleaning_log <- create_cleaning_log(
  raw_data = raw_data, raw_data_uuid = "X_uuid",
  clean_data = clean_data, clean_data_uuid = "X_uuid",
  check_for_deletion_log = T, check_for_variable_name = T
)
cleaning_log
```

### Example:: Comparing cleaning log with clean data and raw data
`review_cleaning_log` function takes raw data, clean data and cleaning log as inputs, and it first creates the cleaning log by comparing raw data and clean data, then compares it with the user-provided cleaning log. Finally, flagged the discrepancies between them (if any). 
``` {r, eval=T, message=FALSE}
compared_df <- review_cleaning_log(
  raw_data = raw_data, raw_data_uuid = "X_uuid",
  clean_data = clean_data, clean_data_uuid = "X_uuid",
  cleaning_log = cleaning_log2, cleaning_log_uuid = "X_uuid",
  cleaning_log_question_name = "questions",
  cleaning_log_new_value = "new_value", cleaning_log_old_value = "old_value",
  deletion_log = deletaion_log, deletion_log_uuid = "X_uuid",
  check_for_deletion_log = T
)

compared_df
```


### Example:: Check of PII
`check_for_pii()` function takes raw data (input can be dataframe or list. However incase of list, you must specify the element name in `element_name` parameter!) and looks for potential PII in the dataset. By default, the function will look for following words but you can also add additional words to look by using `words_to_look` parameter.The default words are-`c("telephone","contact","name","gps","neighbourhood","latitude","logitude","contact","nom","gps","voisinage")`. 
The function will give a list with two element. One will be the data and second one will be the list of potential PII

* Using dataframe as input
``` {r, eval=T}
output_from_data <- check_for_pii(df = raw_data,words_to_look = "date")
output_from_data$potential_PII
```

* Using list as input
``` {r, eval=T}
### from list
df_list <- list(raw_data=raw_data)
output_from_list <- check_for_pii(df = df_list,element_name = "raw_data",words_to_look = "date")
output_from_list$potential_PII
```

### Example:: Check of duration from audits
#### Reading the audits files
It will read only the compressed file.
```{r, eval = FALSE}
my_audit_list <- create_audit_list(audit_zip_path = "audit_for_tests_100.zip")
```

#### Adding the duration to the dataset

Once you have read your audit file from the zip, you will get a list of audit. 
You can use this list to calculate and add the duration. You have 2 options with
a start and end question or summing all the durations.
```{r}
list_audit <- list(uuid1 = data.frame(event = c("form start", rep("question", 5)),
                                      node = c("", paste0("/xx/question", 1:5)),
                                      start = c(1661415887295, 1661415887301,
                                                1661415890819, 1661415892297,
                                                1661415893529, 1661415894720),
                                      end = c(NA, 1661415890790, 1661415892273,
                                              1661415893506, 1661415894703,
                                              1661415896452)),
                   uuid2 = data.frame(event = c("form start", rep("question", 5)),
                                      node = c("", paste0("/xx/question", 1:5)),
                                      start = c(1661415887295, 1661415887301, 1661415890819, 1661415892297, 1661415893529, 1661415894720),
                                      end = c(NA, 1661415890790, 1661415892273, 1661415893506, 1661415894703, 1661415896452)))
some_dataset <- data.frame(X_uuid = c("uuid1", "uuid2"),
                           question1 = c("a","b"),
                           question2 = c("a","b"),
                           question3 = c("a","b"),
                           question4 = c("a","b"),
                           question5 = c("a","b"))
```

If you want to sum all the duration.
```{r}
add_duration_from_audit(some_dataset, uuid_var = "X_uuid", audit_list = list_audit)
```

If you want to use calculate duration between 2 questions.
```{r}
add_duration_from_audit(some_dataset, uuid_var = "X_uuid", audit_list = list_audit,
                        start_question = "question1",
                        end_question = "question3",
                        sum_all = F)
```

If you want to do both.
```{r}
add_duration_from_audit(some_dataset, uuid_var = "X_uuid", audit_list = list_audit,
                        start_question = "question1",
                        end_question = "question3",
                        sum_all = T)
```

#### checking the duration of the dataset
Once you have added the duration to the dataset, you can check if duration are 
between the threshold you are looking for.
``` {r, eval=T}
testdata <- data.frame(
  uuid = c(letters[1:7]),
  duration_audit_start_end_ms = c(2475353, 375491, 2654267, 311585, 817270,
                                  2789505, 8642007),
  duration_audit_start_end_minutes = c(41, 6, 44, 5, 14, 46, 144)
) %>%
  dplyr::rename(`_uuid` = uuid)
check_duration(testdata,
               .col_to_check = "duration_audit_start_end_minutes")
check_duration(
  testdata,
  .col_to_check = "duration_audit_start_end_ms",
  lower_bound = 375490,
  higher_bound = 8642000
)
testdata %>% check_duration(.col_to_check = "duration_audit_start_end_minutes") %>%
  check_duration(
    .col_to_check = "duration_audit_start_end_ms",
    name_log = "duration_in_ms",
    lower_bound = 375490,
    higher_bound = 8642000
  )
```

#### Example:: Check outliers
`check_outliers()` takes raw data set and look for potential outlines. It can both data frame or list. However you must specify the element name (name of your data set in the given list) in `element_name` parameter!  

``` {r, eval=T}
set.seed(122)
### from list
df_outlier<- data.frame(
  uuid = paste0("uuid_", 1:100),
  one_value = c(round(runif(90, min = 45,max =55)), round(runif(5)), round(runif(5,99,100))),
  expense = c(sample(200:500,replace = T,size = 95),c(600,100,80,1020,1050)),
  income = c(c(60,0,80,1020,1050),sample(20000:50000,replace = T,size = 95)),
  yy = c(rep(100,99),10)
)
outliers <- check_outliers(df = df_outlier,uuid_col_name = "uuid")
outliers$potential_outliers

```

#### Example:: Create clean data

`create_clean_data()` function applies cleaning log to the raw data set and returns a clean data set. 

``` {r, eval=T}
cleaning_log_test <- data.frame(
  uuid = paste0("uuid",1:4),
  question= c("age","gender","pop_group","strata"),
  change_type = c("blank_response","no_change","Delete","change_res"),
  new_value = c(NA_character_,NA_character_,NA_character_,"st-a")
)
test_data <- data.frame(
  uuid =  paste0("uuid",1:4),
  age = c(180,23,45,67),
  gender = c("male","female","male","female"),
  pop_group = c("idp","refugee","host","idp"),
  strata = c("a","b","c","d")
)

clean_dataset <- create_clean_data(df = test_data,df_uuid = "uuid",cl = cleaning_log_test,
                                   cl_change_type_col =  "change_type",
                                   values_for_change_response = "change_res",
                                   values_for_blank_response = "blank_response",
                                   values_for_no_change = "no_change",
                                   values_for_remove_survey = "Delete",
                                   cl_change_col =  "question",
                                   cl_uuid = "uuid",
                                   cl_new_val = "new_value" )


```

#### Example:: Check for value

`check_for_value()` function look for specified value in the given data set and return in a cleaning log format. The function can take a data frame or a list as input.

``` {r, eval=T}
set.seed(122)

df <- data.frame(
  X_uuid = paste0("uuid_",1:100),
  age = c(sample(18:80,replace = T,size = 96),99,99,98,88),
  gender = c("99",sample(c("male","female"),replace = T,size = 95),"98","98","88","888"))

output <- check_for_value(df = df,uuid_col_name = "X_uuid",element_name = "checked_dataset",values_to_look = c(99,98,88,888))

output$flaged_value

```



#### Example:: Recreate parent column for choice multiple 

`recreate_parent_column()` recreates the concerted columns for select multiple questions

``` {r, eval=T}
test_data <- dplyr::tibble(
  uuid = paste0("uuid_",1:6),
  gender = rep(c("male","female"),3),
  reason = c("xx,yy","xx,zy",
             "zy","xx,xz,zy",
             NA_character_,"xz"),
  reason.x.x. = c(0,1,0,1,0,0),
  reason.yy = c(1,0,0,0,1,0),
  reason.x.z = c(0,0,0,1,0,1),
  reason.zy = c(0,1,1,1,0,0),
  reason_zy = c(NA_character_,"A","B","C",NA_character_,NA_character_))

recreate_parent_column(df = test_data,uuid = "uuid",sm_sep = ".")

```


#### Example:: Recreate parent column for choice multiple 

`review_sample_frame_with_dataset()` compares the sample frame with dataset and provide the overview of completed and remaining surveys.
``` {r, eval=T,warning =F}
review_output<- review_sample_frame_with_dataset(sample_frame = cleaningtools::sample_frame ,
                                                 sample_frame_strata_col = "Neighbourhood",
                                                 sample_frame_target_survey_col ="Total.no.of.HH",
                                                 clean_data = cleaningtools::clean_data,
                                                 clean_data_strata_column = "neighbourhood",
                                                 consent_column = "consent_remote",
                                                 value_for_consent_yes = "yes")
review_output |> head()
```

#### Example:: Logical checks
```{r}
test_data <- data.frame(uuid = c(1:10) %>% as.character(),
                        today = rep("2023-01-01", 10),
                        location = rep(c("villageA", "villageB"),5),
                        distance_to_market = c(rep("less_30", 5), rep("more_30",5)),
                        access_to_market = c(rep("yes",4), rep("no",6)),
                        number_children_05 = c(rep(c(0,1),4),5,6))
check_for_logical(test_data,
                  uuid_var = "uuid",
                  check_to_perform = "distance_to_market == \"less_30\" & access_to_market == \"no\"",
                   variables_to_clean = "distance_to_market, access_to_market",
                   description = "distance to market less than 30 and no access")
```

```{r}
test_data <- data.frame(uuid = c(1:10) %>% as.character(),
                       distance_to_market = rep(c("less_30","more_30"),5),
                       access_to_market = c(rep("yes",4), rep("no",6)),
                       number_children_05 = c(rep(c(0,1),4),5,6),
                       number_children_618 = c(rep(c(0,1),4),5,6))
check_list <- data.frame(name = c("logical_xx", "logical_yy", "logical_zz"),
                         check = c("distance_to_market == \"less_30\" & access_to_market == \"no\"",
                                   "number_children_05 > 3",
                                   "rowSums(across(starts_with(\"number\")), na.rm = T) > 9"),
                         description = c("distance to market less than 30 and no access",
                                         "number of children under 5 seems high",
                                         "number of children very high"),
                         variables_to_clean = c("distance_to_market, access_to_market",
                                                "number_children_05",
                                                ""))
check_for_logical_with_list(test_data,
                            uuid_var = "uuid",
                            list_of_check = check_list,
                            check_id_column = "name",
                            check_to_perform_column = "check",
                            variables_to_clean_column = "variables_to_clean",
                            description_column = "description")
```
